Title Clustering for recommendations
========================================================
```{r}
movies <- read.table("data/movielens.txt", header=FALSE, sep="|", quote="\"")
colnames(movies) <- c("ID", "Title", "ReleaseDate", "VideoReleaseDate", "IMDB", "Unknown", "Action", "Adventure", "Animation", "Childrens", "Comedy", "Crime", "Documentary", "Drama", "Fantasy", "Filmnoir", "Horror", "Musical", "Mystery", "Romance", "SciFi", "Thriller", "War", "Western")
movies$ID <- NULL
movies$ReleaseDate <- NULL
movies$VideoReleaseDate <- NULL
movies$IMDB <- NULL
movies <- unique(movies)
```
Now we can cluster
```{r}
distances <- dist(movies[, 2:20], method="euclidean")
clusterMovies <- hclust(distances, method="ward")
plot(clusterMovies)
clusterGroups <- cutree(clusterMovies, k = 10)
tapply(movies$Action, clusterGroups, mean)
tapply(movies$Romance, clusterGroups, mean)
clusterLevel10MeanGenre <- data.frame(t(sapply(2:20, function(i) tapply(movies[, i], clusterGroups, mean))))
rownames(clusterLevel10MeanGenre) <- colnames(movies)[2:20]
```
Now recommend, start with an example
```{r}
tit <- "Men in Black (1997)"
c <- clusterGroups[(1:dim(movies)[1])[movies$Title == tit]]
cluster2 <- subset(movies, clusterGroups==c)
```
##Image segmentation
```{r}
flower <- read.csv("data/flower.csv", header=FALSE)
flowerMatrix <- as.matrix(flower)
str(flowerMatrix)
flowerVector <- as.vector(flowerMatrix)
str(flowerVector)
distance <- dist(flowerVector, method="euclidean")
clusterIntensity <- hclust(distance, method="ward")
plot(clusterIntensity)
rect.hclust(clusterIntensity, k=3, border="red")
flowerClusters <- cutree(clusterIntensity, k=3)
tapply(flowerVector, flowerClusters, mean)
dim(flowerClusters) <- c(50,50)
image(flowerClusters, axes=FALSE)
image(flowerMatrix, axes=FALSE, col=grey(seq(0,1,length=256)))
```
Now mri data
```{r}
healthy <- read.csv("data/healthy.csv", header=FALSE)
healthyMatrix <- as.matrix(healthy)
image(healthyMatrix, col=grey(seq(0,1,length=256)))
healthyVector <- as.vector(healthyMatrix)
#kmeans
k <- 5
set.seed(1)
KMC <- kmeans(healthyVector, centers=k, iter.max=1000)
str(KMC)
healthyClusters <- KMC$cluster
KMC$centers
dim(healthyClusters) <- c(nrow(healthyMatrix), ncol(healthyMatrix))
image(healthyClusters, axes=FALSE, col=rainbow(k))
```
Tumors
```{r}
tumor <- read.csv("data/tumor.csv", header=FALSE)
tumorMatrix <- as.matrix(tumor)
tumorVector <- as.vector(tumorMatrix)
install.packages("flexclust")
library(flexclust)
KMC.kcca <- as.kcca(KMC, healthyVector)
tumorClusters <- predict(KMC.kcca, newdata=tumorVector)
dim(tumorClusters) <- c(nrow(tumorMatrix), ncol(tumorMatrix))
image(tumorClusters, axes=FALSE, col=rainbow(k))
```

## Document Clustering
```{r}
dailykos <- read.csv("data/dailykos.csv", stringsAsFactors=FALSE)
distance <- dist(dailykos[, 2:1546], method="euclidean")
dailykosClustering <- hclust(distance, method="ward")
plot(dailykosClustering)
dailykosClusters7 <- cutree(dailykosClustering, k = 7)
dailykosByCluster <- lapply(1:7, function(i)subset(dailykos, dailykosClusters7 == i))
sapply(dailykosByCluster, function(c)dim(c)[1])
lapply(dailykosByCluster, function(c) tail(sort( colMeans(c[-1]))))
```
Lets cluster
```{r}
set.seed(1000)
N <- nrow(dailykos)
KMC <- kmeans(dailykos[, -1],centers=7)
dailykosKmean7 <- lapply(1:7, function(i) subset(dailykos[,-1], KMC$cluster == i))
sapply(dailykosKmean7, function(c) dim(c)[1])
lapply(dailykosKmean7, function(c) tail(sort( colMeans(c[-1]))))
compareKmeansHclust <- sapply(1:7, function(i) sapply(1:7, function(j) sum(dailykosClusters7 == j & KMC$cluster == i) ))
```

##Market Segmentation for Airlines
```{r}
airlines <- read.csv("data/AirlinesCluster.csv")
summary(airlines)
#install.packages("caret")
library(caret)
preproc <- preProcess(airlines)
airlinesNorm <- predict(preproc, airlines)
apply(airlinesNorm, 2, max)
apply(airlinesNorm, 2, min)
distance <- dist(airlinesNorm, method="euclidean")
airlinesClustering <- hclust(distance, method="ward")
plot(airlinesClustering)
airlinesClusters5 <- cutree(airlinesClustering, k=5)
airlinesByCluster <- lapply(1:7, function(i)subset( airlines, airlinesClusters5 == i))
numDataPoints <- sapply(airlinesByCluster, nrow)
clusterMeans <- apply(airlines, 2, function(c) tapply(c, airlinesClusters5, mean))
```
kmeans
```{r}
set.seed(88)
N <- nrow(airlines)
KMC <- kmeans(airlines, centers=5, iter.max=1000)
airlinesKmeansClusters <- lapply(1:5, function(i) subset(airlines, KMC$cluster == i))
sapply( airlinesKmeansClusters, nrow)
kmclusterMeans <- apply(airlines, 2, function(c) tapply(c, KMC$cluster, mean))
