Title
========================================================

```{r}
hotels <- read.csv2("SelectingHotels.csv")
colnames(hotels) <- c("Hotel", "Location", "PriceUn", "Price", "Income", "Students", "Pop")
hotels$PriceUn <- as.numeric(gsub(x=gsub(x=gsub(x=hotels$PriceUn, pattern="$", replacement="", fixed=TRUE), pattern=".", replacement="", fixed=TRUE), pattern=",", replacement=".", fixed=TRUE))

modelParameters <- c(Pop=-5.41, Price=5.86, Income=-3.09, Students=1.75)
modelParameters <- modelParameters[colnames(hotels)[-c(1:3)]]
profitability <- apply(hotels[, -(1:3)], 1, function(x) 39.05 + sum(modelParameters*x))
```

lp model
```{r}
lpHotel <- make.lp(1, nrow(hotels))
constrainMat <- matrix(0, 1, ncol=nrow(hotels))
constrainMat[1, ] <- hotels$PriceUn
for(j in 1:ncol(constrainMat)) set.column(lpHotel, j, constrainMat[, j])
set.rhs(lpHotel, c(1e7) )
set.objfn(lpHotel, profitability)
set.constr.type(lpHotel, c("<="))
lp.control(lpHotel, sense="max")
for(i in 1:nrow(hotels)) set.type(lpHotel, i, "binary")
write.lp(lpHotel, 'hotels.lp', type='lp')
solve(lpHotel)
get.objective(lpHotel)
sum(get.variables(lpHotel))
```
add more constraints
```{r southLakeTahoe}
lpHotel <- make.lp(2, nrow(hotels))
constrainMat <- matrix(0, 2, ncol=nrow(hotels))
constrainMat[1, ] <- hotels$PriceUn
constrainMat[2, 10:16] <- 1
for(j in 1:ncol(constrainMat)) set.column(lpHotel, j, constrainMat[, j])
set.rhs(lpHotel, c(1e7, 2))
set.objfn(lpHotel, profitability)
set.constr.type(lpHotel, c("<="))
lp.control(lpHotel, sense="max")
for(i in 1:nrow(hotels)) set.type(lpHotel, i, "binary")
write.lp(lpHotel, 'hotels.lp', type='lp')
solve(lpHotel)
get.objective(lpHotel)
sum(get.variables(lpHotel))
