Title
========================================================

A trading company is looking for a way to maximixe profit per transportation of their goods.
The company has 3 wagons on a train. They have 4 types of cargo, with a given volume and weight. 
How much of each cargo on which wagon?

Constraints: 
1. weight capacity per train wagon
2. volume capacity per train wagon
3. limited availability per cargo

```{r capacities}
wagonCapacity <- data.frame(weight=c(10, 8, 12), volume=c(5000, 4000, 8000))
cargo <- data.frame(available=c(18, 10, 5, 20), volume=c(400, 300, 200, 500), profit=c(2000, 2500, 5000,3500))
Nw <- nrow(wagonCapacity)
Nc <- nrow(cargo)
```
```{r model}
library(lpSolveAPI)
lpmodel <- make.lp(2*nrow(wagonCapacity) + nrow(cargo), 12)
#constrain variables as a matrix Nw*Nc, coded as a column major vector
#constraints
#1.availability
availIndices <-  lapply(1:Nc, function(j) {1:Nw + Nw*(j-1) })
availValues <- lapply(1:Nc, function(j) rep(1, Nw))
#2. weight
weightIndices <- lapply(1:Nw, function(i) { i + (1:Nc-1)*Nw})
weightValues <- lapply(1:Nw, function(i) rep(1, Nc))
#3. volume
volumeIndices <- weightIndices
volumeValues <-  lapply(1:Nw, function(i) cargo$volume)

constrainIndices <- c(availIndices, weightIndices, volumeIndices)
constrainValues <- c(availValues, weightValues, volumeValues)

constrainMat <- matrix(0, nrow=Nc + 2*Nw, ncol=Nc*Nw)
for( i in 1:(Nc + 2*Nw))  constrainMat[i, constrainIndices[[i]]] <- constrainValues[[i]]

profit <- as.vector(do.call("rbind", lapply(1:Nw, function(i) cargo$profit)))
  
for(j in 1:(Nc*Nw)) set.column(lpmodel, j, constrainMat[, j])
set.objfn(lpmodel, profit)
set.constr.type(lpmodel, rep("<=", 10))               
set.rhs(lpmodel, c(cargo$available, wagonCapacity$weight, wagonCapacity$volume))
lp.control(lpmodel, sense="max")

write.lp(lpmodel, 'wagonCargo.lp', type='lp')
solve(lpmodel)
get.objective(lpmodel)
