---
title: "visualizationRecit"
output: html_document
---
```{r}
library(ggplot2)
intl <- read.csv("~/work/learn/analyticsEdge/data/intl.csv")
intl <- transform(intl, Region=reorder(Region, -PercentOfIntl))
str(intl)
intl$PercentOfIntl <- 100*intl$PercentOfIntl
ggplot(intl, aes(x=Region, y=PercentOfIntl)) +
  geom_bar(stat="identity", fill="dark blue") +
  geom_text(aes(label=PercentOfIntl), vjust=-0.4) + 
  ylab("Percent of International Students") +
  theme(axis.title.x=element_blank(), axis.text.x=element_text(angle=45, hjust=1, size=18), axis.text.y=element_text(size=24), axis.title.y = element_text(size=24))
```
on a world map
```{r worldMap}
library(ggmap)
intlall <- read.csv("data/intlall.csv", stringsAsFactors=FALSE)
head(intlall)
intlall[is.na(intlall)] <- 0
worldMap <- map_data("world")
intlall$Citizenship[ intlall$Citizenship == "China (People's Republic Of)"] <-"China"
worldMap <- merge(worldMap, intlall, by.x="region", by.y="Citizenship")
worldMap <- worldMap[order(worldMap$group, worldMap$order),]
ggplot(worldMap, aes(x=long, y=lat, group=group)) + 
  geom_polygon(aes(fill=Total), color="black") + 
  coord_map("mercator")

ggplot(worldMap, aes(x=long, y=lat, group=group)) + 
  geom_polygon(aes(fill=Total), color="black") + 
  coord_map("ortho", orientation=c(-37, 30, 0))
