---
title: "Visual Homework"
output: html_document
---
##Election Forecast
```{r electionForecast}
library(ggplot2)
library(maps)
library(ggmap)
statesMap <- map_data("state")
table(statesMap$group)
ggplot(statesMap, aes(x=long, y=lat, group=group)) + 
  geom_polygon(fill="white", color="red") +
  coord_map("mercator")

polling <- read.csv("data/PollingImputed.csv")
train <- subset(polling, Year < 2012)
test <- subset(polling, Year == 2012)
mod2 <- glm(Republican ~ SurveyUSA + DiffCount, data=train, family="binomial")
testPred <- predict(mod2, newdata=test, type="response")
testPredBinary <- as.numeric(testPred > 0.5)
predDataFrame <- data.frame(testPred, testPredBinary, test$State)
predDataFrame$region <- tolower(predDataFrame$test.State)
predMap <- merge(statesMap, predDataFrame, by="region")
predMap <- predMap[order(predMap$order),]

ggplot(predMap, aes(x=long, y=lat, group=group, fill=testPredBinary)) + 
  geom_polygon(color="black") + 
  scale_fill_gradient(low="blue", high="red", guide="legend", breaks=c(0,1), labels=c("Democrate", "Republican"), name="Prediction 2012")

ggplot(predMap, aes(x=long, y=lat, group=group, fill=testPred)) + 
  geom_polygon(color="black", size=3, alpha = 0.3)+ 
  scale_fill_gradient(low="blue", high="red", guide="legend", breaks=c(0,1), labels=c("Democrate", "Republican"), name="Prediction 2012")

```
##Paroles

```{r parole}
parole <- read.csv("data/parole.csv")
parole$male <- as.factor(parole$male)
parole$state <- as.factor(parole$state)
parole$crime <- as.factor(parole$crime)
table(parole$male)

ggplot(data=parole, aes(x=age)) + 
  geom_histogram(binwidth=5, color="blue")

ggplot(data=parole, aes(x=age)) +
  geom_histogram(binwidth=5, color="white") + facet_grid(male~.)

ggplot(data=parole, aes(x=age)) +
  geom_histogram(binwidth=5, color="white") + facet_grid(. ~ male)

ggplot(data=parole, aes(x=age, fill=male)) +
  geom_histogram(binwidth=5, color="white")

ggplot(data=parole, aes(x=age, fill=male)) +
  geom_histogram(binwidth=5, position="identity", alpha = 0.5, color="white")

ggplot(data=parole, aes(x=time.served)) + 
  geom_histogram(binwidth=1, color="white") + facet_grid(crime ~ .)

ggplot(data=parole, aes(x=time.served, fill=crime)) + 
  geom_histogram(binwidth=1, position="identity",alpha=0.5, color="white")

```
Network

```{r network}
edges <- read.csv("data/edges.csv")
users <- read.csv("data/users.csv")
library(igraph)
g <- graph.data.frame(edges, FALSE, users)
plot(g, vertex.size=5, vertex.label=NA)
V(g)$size <- degree(g)/2 + 2
plot(g, vertex.label=NA)
max(V(g)$size)
min(V(g)$size)
V(g)$color = "black"
V(g)$color[V(g)$gender == "A"] = "red"
V(g)$color[V(g)$gender == "B"] = "gray"
plot(g, vertex.label=NA)
V(g)$color = "black"
V(g)$color[V(g)$school == "A"] = "red"
plot(g, vertex.label=NA)
V(g)$color = "black"
V(g)$color[V(g)$locale == "A"] = "red"
V(g)$color[V(g)$locale == "B"] = "green"
rglplot(g, vertex.label=NA, edge.width=2)
```

Tweets

```{r tweets}
library(tm)
library(SnowballC)
library(wordcloud)
tweets = read.csv("data/tweets.csv", stringsAsFactors=FALSE)
corpus <- Corpus(VectorSource(tweets$Tweet))
corpus <- tm_map(corpus, tolower)
corpus <- tm_map(corpus, removePunctuation)
corpus <- tm_map(corpus, removeWords, c("apple", stopwords("english")))
frequencies <- DocumentTermMatrix(corpus)
allTweets <- as.data.frame(as.matrix(frequencies))

wordcloud(colnames(allTweets), colSums(allTweets))
wordcloud(colnames(allTweets[tweets$Avg < 0, ]), colSums(allTweets[tweets$Avg < 0,]), scale=c(4, 0.5), random.order = FALSE, min.freq = 5, colors=brewer.pal(9, "Blues")[c(5,6,7,8,9)])
