---
title: "Interacting Variables"
output: html_document
---
Time now to play with interactions
```{r dataIn}
compData <- read.csv("train.csv")
questions <- colnames(compData)[9:109]
library(caTools)
split <- sample.split( compData$Happy, SplitRatio = 0.7)
compTrain <- subset(compData, split==TRUE)
compTest <- subset(compData, split==FALSE)
```
Levels in the data, 
```{r qlevels}

qlevels <- t(sapply(questions, function(q) levels(compData[,q])))
rownames(qlevels) <- questions
hlevels <- levels(compData$HouseholdStatus)

```

```{r natureOfTheQuestions}
qidx <- read.csv("questions.csv", stringsAsFactors=FALSE)$Index
qs <- read.csv("questions.csv", stringsAsFactors=FALSE)$Question
qs <- gsub(pattern="/", replacement=" ", qs)
qs <- gsub(pattern="-", replacement=" ", qs)
library(tm)
library(SnowballC)
corpus <- Corpus(VectorSource(qs))
corpus <- tm_map(corpus, tolower)
corpus <- tm_map(corpus, removePunctuation)
corpus <- tm_map(corpus, removeWords, stopwords("english") ) 
capFirstLetter <- function(x) paste(toupper(substring(x, 1,1)), substring(x, 2), sep="")

qlists <- lapply(corpus, function(x) Filter(function(s) s != "", unlist(strsplit(x, " ", fixed=TRUE))))
qsAbrv <- unlist(lapply(qlists, function(xs) paste(Map(capFirstLetter, xs), collapse = "") ))

choices <- read.csv("questions.csv", stringsAsFactors=FALSE)$Choices
qsans <- as.matrix(cbind( qsAbrv, choices))
```
Some useful code
```{r groupStatsByLevel}
happyMean <- function(comp) mean(comp$Happy)
computeGroupMeanByLevel <- function(q, comp){
  m <- sapply(levels( comp[ , q]), 
           function(l) {
             s <- sum( comp$Happy[ comp[, q] == l])
             n <- sum(comp[,q]==l)
             if (s==0) 0 else (s/n)
          }
      )
  m
}
computeGroupSdByLevel <- function(q, comp){
  s <-  sapply(levels( comp[ , q]), 
           function(l) {
             s <- sum( comp$Happy[ comp[, q] == l])
             if (s==0) 0 else sd( comp$Happy[comp[,q] == l])
          }
        )
  s
}
computeGroupSizesByLevel<- function(q, comp){
  s <- sapply(levels( comp[ , q]), function(l) sum( comp[,q] == l))
  s
}
computeGroupMeanSdsByLevel <- function(q, comp) {
  sqrt(2*computeGroupMeanByLevel(q, comp)/computeGroupSizesByLevel(q, comp))
}


energyFromMean <- function(m) -log(m/(1-m))

probFromEnergies <- function(es, hmean) {
  e0 <- -log(hmean/(1-hmean))
  totE <- sum( es - e0) + e0
  1/(1 + exp(totE))
}

computeGroupEnergyByLevel <- function(q, comp){
  energyFromMean( computeGroupMeanByLevel(q, comp))
}
```
To compute total energy of a state-vector, we will need access to the energy levels in each state component. Because the state components do not have the same number of levels, we will have to represent their energies in a list carrying the variable names. Lets start by extracting level statistics
```{r statDefs}
hmean <- happyMean(compTrain)
groupMeans.1 <- lapply(c(questions, "HouseholdStatus"), function(q) computeGroupMeanByLevel(q, compTrain))
names(groupMeans.1) <- c(questions, "HouseholdStatus")
groupSizes.1 <- lapply(c(questions, "HouseholdStatus"), function(q) computeGroupSizesByLevel(q, compTrain))
names(groupSizes.1) <- c(questions, "HouseholdStatus")
groupMeanSds.1 <- lapply(c(questions, "HouseholdStatus"),  function(q) computeGroupMeanSdsByLevel(q, compTrain))
names(groupMeanSds.1) <- c(questions, "HouseholdStatus")
groupEnergies.1 <- lapply(groupMeans.1, function(m) energyFromMean(m))
groupZs.1 <- lapply(1:length(groupMeans.1), function(i) (groupMeans.1[[i]] - hmean)/groupMeanSds.1[[i]])
names(groupZs.1) <- c(questions, "HouseholdStatus")
vs.1 <- c(questions, "HouseholdStatus")
```
Lets order the groupEnergies to determine which one is the best discriminant
```{r svord}
singleVariables <- c(questions, "HouseholdStatus")
ge <- groupEnergies.1
spread <- function(es) max(es) - min(es)
svord <- singleVariables[order(sapply(ge, spread), decreasing=TRUE)]
```
```

These are used to compute the total energy
```{r totalEnergyAndProbHappy}

totalEnergy <- function(anss, ge, hmean){ #anss is a vector of factor variables
  qns <- names(anss)
  levs <- as.numeric(anss)
  sum(sapply(names(anss), function(q) ge[[q]][anss[[q]]])) - log(hmean/(1-hmean))
}

probHappy <- function(anss, ge, hmean) 1/(1+exp(totalEnergy(anss, ge, hmean)))
```
Lets write a function to run regression using energies
```{r energyLinearModel}
elm <- function(response, predictors, data){
  model <- list()
  model$response <- response
  model$predictors <- predictors
  model$data <- data
  R <- data[, response]
  Qs <- as.data.frame(data[, predictors])
  colnames(Qs) <- predictors
  statistics <- list()
  statistics$N <- nrow(Qs)
  statistics$M <- ncol(Qs)
  statistics$ymean <- mean(R)
  statistics$groupMeans <- lapply(Qs, function(x){
    sapply(levels(x), function(l){
       s <- sum( R[ x == l])
       n <- sum(x==l)
       if (s==0) 0 else (s/n)
    })
  })  
  statistics$groupSizes <- lapply(Qs, function(x){
    sapply(levels(x), function(l){
      sum(x==l)
    })
  })
  statistics$groupMeanSds <- lapply(1:statistics$M, function(i){
    sqrt(2*statistics$groupMeans[[i]]/statistics$groupSizes[[i]])
  })
  names(statistics$groupMeanSds) <- predictors
  
  statistics$groupZs <- lapply(1:statistics$M, function(i){
    (statistics$groupMeans[[i]] - statistics$ymean)/statistics$groupMeanSds[[i]]
  })
  names(statistics$groupZs) <- predictors
  
  statistics$groupEnergies <- lapply(statistics$groupMeans, function(m) -log(m/(1-m)) )
  model$statistics <- statistics
  model
}

predict.elm <- function(model, newdata){
  tes <- do.call("cbind", lapply( model$predictors, function(q){
    model$statistics$groupEnergies[[q]][newdata[, q]]
  }))
  rownames(tes) <- rownames(newdata)
  apply(tes, 1, function(es)probFromEnergies(es, model$statistics$ymean))
}
```




###Interactions
```{r}
interactionVariable <- function(u, v){
  if(is.numeric(u) & is.numeric(v)) u + v
  else {if (is.factor(u) & is.factor(v)) sapply(1:length(u), function(i) paste(as.character(u[i]), as.character(v[i]), sep="."))
  else print("Incompatible variable types")
  }
}

makeInteraction <- function(cd){ # assuming two columns of factors
  uv <- list(sapply(1:nrow(cd), function(i) paste(as.character(cd[i,1]), as.character(cd[i,2]), sep=".")))
  names(uv) <- paste(colnames(cd)[1], colnames(cd)[2], sep=".")
  as.data.frame(uv)
}
  
  
varsToInteract <- c(svord[1:2], svord[4:8])
n <- length(varsToInteract)
compDataIntVars <- data.frame(do.call("cbind", lapply(1:(n-1), function(i){ do.call("cbind", lapply((i+1):n, function(j) interactionVariable(compData[, varsToInteract[i]], compData[, varsToInteract[j]]))))))
ivs.2 <-  do.call("c", c(sapply( 1:(n-1), function(i) sapply((i+1):n, function(j) paste(varsToInteract[i], varsToInteract[j], sep="."))) ))
colnames(compDataIntVars) <- ivs.2

compDataIntVars <- lapply(1:(n-1), function(i) do.call("cbind", lapply((i+1):n, function(j) makeInteraction(compData[, varsToInteract[c(i, j)]]))))
compDataIntVars <- data.frame( do.call("cbind", compDataIntVars))
compDataIntVars$Happy <- compData$Happy

cdivs.train <- subset(compDataIntVars, split==TRUE)
cdivs.test <- subset(compDataIntVars, split==FALSE)

cd.train <- cbind(cdivs.train, compTrain[, c(varsToInteract, "HouseholdStatus")])
cd.test <- cbind(cdivs.test, compTest[, c(varsToInteract, "HouseholdStatus")])
ivs <- c(ivs.2, c(varsToInteract, "HouseholdStatus"))
```


Let us do some analysis on the interacting variables.
```{r}
hmean <- happyMean(cd.train)
groupMeans <- lapply(ivs, function(q) computeGroupMeanByLevel(q, cd.train))
names(groupMeans) <- ivs
groupSizes <- lapply(ivs, function(q) computeGroupSizesByLevel(q, cd.train))
names(groupSizes) <- ivs
groupMeanSds <- lapply(ivs, function(q) computeGroupMeanSdsByLevel(q, cd.train))
names(groupMeanSds) <- ivs
groupEnergies <- lapply(groupMeans, function(m) energyFromMean(m, hmean))
groupZs <- lapply(1:length(groupMeans), function(i) (groupMeans[[i]] - hmean)/groupMeanSds[[i]])
names(groupZs) <- ivs


bivStats <- function(i){
  m <- which.max(groupZs[[i]])
  c(sum(groupMeans[[i]][(1:9)[-5]]*groupSizes[[i]][(1:9)[-5]])/sum(groupSizes[[i]][(1:9)[-5]]), groupMeans[[i]][m], groupMeanSds[[i]][m])
}
makeBivariate <- function(i,  vs, gZs, cd){
  m <- which.max(abs(gZs[[vs[i]]]))
  nv <- sapply(cd[, vs[i]], function(x) (x == names(m)))
  as.factor(nv)
}

cdbv.train <- data.frame(do.call("cbind", lapply(1:21, function(i) makeBivariate(i, cdivs.train))))
cdbv.train$Happy <- cdivs.train$Happy
cdbv.test <- data.frame(do.call("cbind", lapply(1:21, function(i) makeBivariate(i, cdivs.test))))
cdbv.test$Happy <- cdivs.test$Happy
cdbv.train <- cbind(cdbv.train, compTrain[, c(varsToInteract, "HouseholdStatus")])
cdbv.test <- cbind(cdbv.test, compTest[, c(varsToInteract, "HouseholdStatus")])
              
  
  
  
                                             
```

Correlations, mutual information
```{r}
probLevels.singleVariable <- function(q, cd){
  sapply(levels(cd[,q]), function(l) sum(cd[, q] == l))/nrow(cd)
}
entropy <- function(m, base=2) -sum(m*log(m)/log(base))

computeMutualInformation <- function(varsToInteract, cd.train){
  n <- length(varsToInteract)
  h1s <- sapply(varsToInteract, function(v) entropy(probLevels.singleVariable(v, cd.train),3))
  mis <- sapply(1:(n-1), function(i) {
      v1 <- varsToInteract[i]
      sapply((i+1):n, function(j){
      v2 <- varsToInteract[j]
      v12 <- paste(v1, v2, sep=".")
      H12 <- entropy(probLevels.singleVariable( v12, cd.train),3)
      H1 <- entropy(probLevels.singleVariable(v1, cd.train),3)
      H2 <- entropy(probLevels.singleVariable(v2, cd.train), 3)
      H1 + H2 - H12
    })
  })
  mismat <- matrix(0, nrow=n, ncol=n)
  for(i in 1:(n-1)){
    mismat[i,i] <- h1s[i]
    for(j in (i+1):n){
      mismat[i,j] <- mis[[i]][j-i]
      mismat[j,i] <- mismat[i,j]
    }
  }
  mismat[n,n] <- h1s[n]
  mismat
}
  

Warm up on the single variables
```{r}
tabacc <- function(tb) sum(diag(tb))/sum(tb)
ivs1 <- c(svord[1:5]) #, "HouseholdStatus")
prob1.train <- sapply(1:nrow(compTrain), function(i) probHappy(compTrain[i, ivs1], ge=groupEnergies.1, hmean=hmean))
prob1.test <- sapply(1:nrow(compTest), function(i) probHappy(compTest[i, ivs1], ge=groupEnergies.1, hmean=hmean))
tb.train <- table(compTrain$Happy, prob1.train > 0.5)
tb.test <- table(compTest$Happy, prob1.test > 0.5)
tabacc(tb.train)
tabacc(tb.test)
```
For interacting variables
```{r}
prob.train <- sapply(1:nrow(cd.train), function(i) probHappy(cd.train[i, ivs], ge=groupEnergies, hmean=hmean))
prob.test <- sapply(1:nrow(cd.test ), function(i) probHappy(cd.test [i, ivs], ge=groupEnergies, hmean=hmean))
tb.train <- table(cdivs.train$Happy, prob.train > 0.5)
tb.test <- table(cdivs.test$Happy, prob.test > 0.5)
tabacc(tb.train)
tabacc(tb.test)
```
Tree
```{r}
cd2.tree <- rpart( Happy ~ ., cd.train, method="class")
cd2.tree.pred <- predict(cd2.tree, newdata=cd.test, type="class")
tabacc( table(cd.test$Happy, cd2.tree.pred ))
                     
```

Plot means and sds

```{r}
n <- 1
cols <- c("orange1", "orange2", "orange3", "orange4", "red1", "red2", "red3", "red4", "orangered")
cols <- c("red", "green", "blue", "orange", "yellow", "magenta", "brown", "cyan", "black")
plot(groupMeans[[n]], col=cols, pch=19, ylim=c(0,1))
abline(h = groupMeans[[n]] - groupMeanSds[[n]], col=cols)
abline(h = groupMeans[[n]] + groupMeanSds[[n]], col=cols)


###Are the questions independent?

```{r}
iv.gems.1 <- do.call("rbind", lapply( 11:15, function(i) sapply(levels(ivs.train[,i]), function(l) mean( ivs.train[ ivs.train[,i] == l, 16])/h)))
iv.Pgh.1 <- iv.gems.1/rowSums(iv.gems.1)
iv.lPgh.1 <- -log(iv.Pgh.1)
rownames(iv.lPgh.1) <- colnames(ivs.train)[11:15]

iv.gems.2 <- do.call("rbind", lapply( 1:10, function(i) sapply(levels(ivs.train[,i]), function(l) mean( ivs.train[ ivs.train[,i] == l, 16])/h)))
iv.Pgh.2 <- iv.gems.2/rowSums(iv.gems.2)
iv.lPgh.2 <- -log(iv.Pgh.2)
rownames(iv.lPgh.2) <- colnames(ivs.train)[1:10]

indepLogProb <- function(i, j){
  e <- c(sapply(1:3, function(l) sapply(1:3, function(m) iv.lPgh.1[i,l] + iv.lPgh.1[j,m])))
  names(e) <- c(sapply(levels(comp[,rownames(iv.lPgh.1)[i]]), function(l) sapply( levels(comp[, rownames(iv.lPgh.1)[j]]), function(m) paste(l,m))))
  e
}
iv.lPgh.2.ind <- do.call("rbind", lapply(1:4, function(i) do.call("rbind", lapply((i+1):5, function(j) indepLogProb(i,j)))))

entropy <- function(p) -sum(p*log(p)/log(2))

mi <- sapply(1:10, function(i) entropy( exp(-iv.lPgh.2[i, ])) - entropy(exp(-iv.lPgh.2.ind[i,])))

iv.Phg.1 <-  do.call("rbind", lapply( 11:15, function(i) sapply(levels(ivs.train[,i]), function(l) mean( ivs.train[ ivs.train[,i] == l, 16]))))
rownames(iv.Phg.1) <- colnames(ivs.train)[11:15]

iv.Phg.2 <- do.call("rbind", lapply( 1:10, function(i) sapply(levels(ivs.train[,i]), function(l) mean( ivs.train[ ivs.train[,i] == l, 16]))))
rownames(iv.Phg.2) <- colnames(ivs.train)[1:10]

e12 <- t(apply(iv.Phg.2, 1, function(m) -log(m/(1-m))))
rownames(e12) <- colnames(ivs.train)[1:10]
```



```
